---
title: "RWorksheet_rocillo#4a"
author: "Cassandra Jielin Rocillo"
date: "2024-10-17"
output:
  pdf_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
1. The table below shows the data about shoe size and height. Create a data frame.

```{r}
  shoe_size <- c(6.5, 9.0, 8.5, 8.5, 10.5, 7.0, 9.5, 9.0, 13.0, 7.5, 10.5, 8.5, 12.0, 10.5, 13.0, 11.5, 8.5, 5.0, 10.0, 6.5, 7.5, 8.5, 10.5, 8.5, 10.5, 11.0, 9.0, 13.0)
  height <- c(66.0, 68.0, 64.5, 65.0, 70.0, 64.0, 70.0, 71.0, 72.0, 64.0, 74.5, 67.0, 71.0, 71.0, 77.0, 72.0, 59.0, 62.0, 72.0, 66.0, 64.0, 67.0, 73.0, 69.0, 72.0, 70.0, 69.0, 70.0)
  gender <- c("F", "F", "F", "F", "M", "F", "F", "F", "M", "F", "M", "F", "M", "M", "M", "M", "F", "F", "M", "F", "F", "M", "M", "F", "M", "M", "M", "M")
  
data <- data.frame(Shoe_size = shoe_size, Height = height, Gender = gender)

```

a. Describe the data.
  - It shows the shoe size, height and gender.
 
b. Create a subset by males and females with their corresponding shoe size and height.
What its result? Show the R scripts.

```{r}
male_subset <- subset(data, gender == "M")
female_subset <- subset(data, gender == "F")

print(male_subset)
print(female_subset)

print(data)
```

c. Find the mean of shoe size and height of the respondents. Write the R scripts and its
result.

```{r}
  mean_of_shoe_size <- mean(shoe_size)
  mean_of_height <- mean(height)
  
  shoe_size_mean <- paste("Mean of shoe size",mean_of_shoe_size)
  height_mean <- paste("Mean of height", mean_of_height)
  
  print(shoe_size_mean)
  print(height_mean)
  
```
d. Is there a relationship between shoe size and height? Why?
  Yes, there is a relationship between shoe size and height, as taller individuals often have larger feet. 
  
2. Construct character vector months to a factor with factor() and assign the result to factor_months_vector.
Print out factor_months_vector and assert that R prints out the factor levels below the actual values.
Consider data consisting of the names of months:
("March", "April", "January", "November", "January", "September", "October", "September", "November", "August", "January", "November", "November", "February", "May", "August", "July", "December", "August", "August", "September", "November", "February", "April")

```{r}
  months_vector <- c("March", "April", "January", "November", "January", "September", "October", "September", "November", "August", "January", "November", "November", "February", "May", "August", "July", "December", "August", "August", "September", "November", "February", "April")

  factor_months_vector <- factor(months_vector)
  factor_months_vector
```
3. Then check the summary() of the months_vector and factor_months_vector.
Interpret the results of both vectors.

```{r}
 summary(months_vector)

 summary(factor_months_vector)
 
 
```
Are they both equally useful in this case?
  - the factor_months_vector is more useful because it shows specifically what's on your data and it categorized the months and provides the count of how many times each month repeats. 


4. Create a vector and factor for the table below.

Note: Apply the factor function with required order of the level.
new_order_data <- factor(factor_data,levels = c("East","West","North"))
print(new_order_data)

```{r}
  direction <- c("East", "West", "North")
  frequency <- c(1, 4, 3)
  
  data <- data.frame(direction, frequency)
  data
  
  new_order_data <- factor(direction,levels = c("East","West","North"))
print(new_order_data)

```
5. 5. Enter the data below in Excel with file name = import_march.csv

a. Import the excel file into the Environment Pane using read.table() function.
Write the code. 

```{r}
   library(readxl)
data <- read.table("/cloud/project/worksheet/worksheet4/import_march.csv", header = TRUE, sep = ",")
```

b. View the dataset. Write the R scripts and its result.

```{r}
  print(data)
```

