---
title: "RWorksheet_rocillo#4b"
author: "Cassandra Jielin Rocillo"
date: "2024-10-30"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

for() loop

1. Using the for loop, create an R script that will display a 5x5 matrix as shown in

Figure 1. It must contain vectorA = [1,2,3,4,5] and a 5 x 5 zero matrix.
Hint Use abs() function to get the absolute value

```{r}
vectorA <- c(1,2,3,4,5)
matrixx <- matrix(0, nrow = 5, ncol = 5)

for(i in 1:5) {
 for(j in 1:5){
   matrixx[i, j] <- abs(vectorA[i] - vectorA[j])
}
}
print(matrixx)
```

2. Print the string "*" using for() function. The output should be the same as shown

in Figure 2.

```{r}
for(i in 1:5){
  cat(rep('"*"', i), "\n")
}
```

3.Get an input from the user to print the Fibonacci sequence starting from the 1st input

up to 500. Use repeat and break statements. Write the R Scripts and its output

```{r}
x <- 0
y <- 1

z <- readline(prompt = "Enter a number: ")

repeat {
  z <- x + y 
  if (z > 500) break
  x <- y
  y <- z
  print(z)
}
```

4. Import the dataset as shown in Figure 1 you have created previously.

a. What is the R script for importing an excel or a csv file? Display the first 6 rows of the dataset? Show your codes and its result.

```{r}
 library(readr)
shoe_size <- read.csv(file = "ShoeSizes.csv",header = T, sep = ",")

head(shoe_size)
```

b. Create a subset for gender(female and male). How many observations are there in Male? How about in Female? Write the R scripts and its output.

```{r}
female_data <- subset(shoe_size, Gender == "F")
male_data <- subset(shoe_size, Gender == "M")

num_females <- nrow(female_data)
num_males <- nrow(male_data)

num_females
num_males
```
c. Create a graph for the number of males and females for Household Data. Use plot(), chart type = barplot. Make sure to place title, legends, and colors. 
Write the R scripts and its result.

```{r}
  library(ggplot2)
qplot(Gender, data = shoe_size,
geom = "bar",
fill = Gender)
```

5. The monthly income of Dela Cruz family was spent on the following:
  a. Create a piechart that will include labels in percentage.Add some colors and title of the chart. Write the R scripts and show its output.
  
```{r}
categories <- c("Food", "Electricity", "Savings", "Miscellaneous")
expenses <- c(60, 10, 5, 25)

percentages <- round(expenses / sum(expenses) * 100)

pie(expenses, labels = paste(categories, percentages, "%"), 
    col = rainbow(length(categories)),
    main = "Dela Cruz Family Monthly Income Distribution")
```

6. Use the iris dataset.
data(iris)
a. Check for the structure of the dataset using the str() function. 

```{r}
 data(iris)
str(iris)
```

Describe what you have seen in the output.
  - The output shows a data frame with 5 variable sepal.lenght, width, petal.lenght, width and species and the values of the variables.
  
b. Create an R object that will contain the mean of the sepal.length, sepal.width,petal.length,and petal.width. What is the R script and its result?

```{r}
 mean_values <- colMeans(iris[, 1:4])
 mean_values

```
c. Create a pie chart for the Species distribution. Add title, legends, and colors. Write the R script and its result.

```{r}
  species_counts <- table(iris$Species)
pie(species_counts,
    labels = paste(names(species_counts), round(species_counts / sum(species_counts) * 100, 1), "%"),
    col = rainbow(length(species_counts)),
    main = "Species Distribution in Iris Dataset")
```

d. Subset the species into setosa, versicolor, and virginica. Write the R scripts and show the last six (6) rows of each species.

```{r}
setosa <- subset(iris, Species == "setosa")
versicolor <- subset(iris, Species == "versicolor")
virginica <- subset(iris, Species == "virginica")

tail(setosa)
tail(versicolor)
tail(virginica)
```
e. Create a scatterplot of the sepal.length and sepal.width using the different species(setosa,versicolor,virginica). 
Add a title = “Iris Dataset”, subtitle = “Sepal width and length, labels for the x and y axis, the pch symbol and colors should be based on the species.

```{r}
data_iris <- iris
data_iris %>%
ggplot(aes(x = Sepal.Length, y = Sepal.Width, group = Species, color = Species)) +
ggtitle("Iris Data", subtitle = "Sepal width and length") +
geom_point()
```

f. Interpret the result.
  - the data shows the Iris Data and its 3 different species and the sepal width and sepal length of the Iris.
  
7. Import the alexa-file.xlsx. Check on the variations. 
Notice that there are extra whitespaces among black variants (Black Dot, Black Plus, Black Show, Black Spot). Also on the white variants (White Dot, White Plus, White Show, White Spot).