---
title: "RWorksheet_rocillo#4b"
author: "Cassandra Jielin Rocillo"
date: "2024-10-30"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

for() loop

1. Using the for loop, create an R script that will display a 5x5 matrix as shown in

Figure 1. It must contain vectorA = [1,2,3,4,5] and a 5 x 5 zero matrix.
Hint Use abs() function to get the absolute value

```{r}
vectorA <- c(1,2,3,4,5)
matrixx <- matrix(0, nrow = 5, ncol = 5)

for(i in 1:5) {
 for(j in 1:5){
   matrixx[i, j] <- abs(vectorA[i] - vectorA[j])
}
}
print(matrixx)
```

2. Print the string "*" using for() function. The output should be the same as shown

in Figure 2.

```{r}
for(i in 1:5){
  cat(rep('"*"', i), "\n")
}
```

3.Get an input from the user to print the Fibonacci sequence starting from the 1st input

up to 500. Use repeat and break statements. Write the R Scripts and its output

```{r}
x <- 0
y <- 1

z <- readline(prompt = "Enter a number: ")

repeat {
  z <- x + y 
  if (z > 500) break
  x <- y
  y <- z
  print(z)
}
```

4. Import the dataset as shown in Figure 1 you have created previously.

a. What is the R script for importing an excel or a csv file? Display the first 6 rows of the dataset? Show your codes and its result.

```{r}
 library(readr)
shoe_size <- read.csv(file = "ShoeSizes.csv",header = T, sep = ",")

head(shoe_size)
```

b. Create a subset for gender(female and male). How many observations are there in Male? How about in Female? Write the R scripts and its output.

```{r}
female_data <- subset(shoe_size, Gender == "F")
male_data <- subset(shoe_size, Gender == "M")

num_females <- nrow(female_data)
num_males <- nrow(male_data)

num_females
num_males
```
c. Create a graph for the number of males and females for Household Data. Use plot(), chart type = barplot. Make sure to place title, legends, and colors. 
Write the R scripts and its result.

```{r}
  library(ggplot2)
qplot(Gender, data = shoe_size,
geom = "bar",
fill = Gender)
```

5. The monthly income of Dela Cruz family was spent on the following:
  a. Create a piechart that will include labels in percentage.Add some colors and title of the chart. Write the R scripts and show its output.
  



