---
title: "RWorksheet_rocillo#4a"
author: "Cassandra Jielin Rocillo"
date: "2024-10-17"
output:
  pdf_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
1. The table below shows the data about shoe size and height. Create a data frame.

```{r}
  shoe_size <- c(6.5, 9.0, 8.5, 8.5, 10.5, 7.0, 9.5, 9.0, 13.0, 7.5, 10.5, 8.5, 12.0, 10.5, 13.0, 11.5, 8.5, 5.0, 10.0, 6.5, 7.5, 8.5, 10.5, 8.5, 10.5, 11.0, 9.0, 13.0)
  height <- c(66.0, 68.0, 64.5, 65.0, 70.0, 64.0, 70.0, 71.0, 72.0, 64.0, 74.5, 67.0, 71.0, 71.0, 77.0, 72.0, 59.0, 62.0, 72.0, 66.0, 64.0, 67.0, 73.0, 69.0, 72.0, 70.0, 69.0, 70.0)
  gender <- c("F", "F", "F", "F", "M", "F", "F", "F", "M", "F", "M", "F", "M", "M", "M", "M", "F", "F", "M", "F", "F", "M", "M", "F", "M", "M", "M", "M")
  
data <- data.frame(Shoe_size = shoe_size, Height = height, Gender = gender)

```

a. Describe the data.
  - It shows the shoe size, height and gender.
 
b. Create a subset by males and females with their corresponding shoe size and height.
What its result? Show the R scripts.

```{r}
male_subset <- subset(data, gender == "M")
female_subset <- subset(data, gender == "F")

print(male_subset)
print(female_subset)

print(data)
```

c. Find the mean of shoe size and height of the respondents. Write the R scripts and its
result.

```{r}
  mean_of_shoe_size <- mean(shoe_size)
  mean_of_height <- mean(height)
  
  shoe_size_mean <- paste("Mean of shoe size",mean_of_shoe_size)
  height_mean <- paste("Mean of height", mean_of_height)
  
  print(shoe_size_mean)
  print(height_mean)
  
```
d. Is there a relationship between shoe size and height? Why?
  Yes, there is a relationship between shoe size and height, as taller individuals often have larger feet. 
  
2. Construct character vector months to a factor with factor() and assign the result to factor_months_vector.
Print out factor_months_vector and assert that R prints out the factor levels below the actual values.
Consider data consisting of the names of months:
("March", "April", "January", "November", "January", "September", "October", "September", "November", "August", "January", "November", "November", "February", "May", "August", "July", "December", "August", "August", "September", "November", "February", "April")

```{r}
  months_vector <- c("March", "April", "January", "November", "January", "September", "October", "September", "November", "August", "January", "November", "November", "February", "May", "August", "July", "December", "August", "August", "September", "November", "February", "April")

  factor_months_vector <- factor(months_vector)
  factor_months_vector
```
3. Then check the summary() of the months_vector and factor_months_vector.
Interpret the results of both vectors.

```{r}
 summary(months_vector)

 summary(factor_months_vector)
 
 
```
Are they both equally useful in this case?
  - the factor_months_vector is more useful because it shows specifically what's on your data and it categorized the months and provides the count of how many times each month repeats. 


4. Create a vector and factor for the table below.

Note: Apply the factor function with required order of the level.
new_order_data <- factor(factor_data,levels = c("East","West","North"))
print(new_order_data)

```{r}
  direction <- c("East", "West", "North")
  frequency <- c(1, 4, 3)
  
  data <- data.frame(direction, frequency)
  data
  
  new_order_data <- factor(direction,levels = c("East","West","North"))
print(new_order_data)

```
5. 5. Enter the data below in Excel with file name = import_march.csv

a. Import the excel file into the Environment Pane using read.table() function.
Write the code. 

```{r}
   library(readxl)
data <- read.table("/cloud/project/worksheet/worksheet 4/import_march.csv", header = TRUE, sep = ",")
```

b. View the dataset. Write the R scripts and its result.

```{r}
  print(data)
```

6. Full Search
  Exhaustive search is a methodology for finding an answer by exploring all possible cases.
When trying to find a desired number in a set of given numbers, the method of finding the corresponding number by checking all elements in the set one by one can be called an exhaustive search. Implement an exhaustive search function that meets the input/output conditions below.

a. Create an R Program that allows the User to randomly select numbers from 1 to 50.
Then display the chosen number. If the number is beyond the range of the selected choice, it will have to display a string “The number selected is beyond the range of 1 to 50”. 
If number 20 is inputted by the User, it will have to display “TRUE”, otherwise display the input number.

```{r}
num <- readline(prompt = "Enter a number: ")
num 

if(num < 1 || num > 50) {
  paste("The number selected is beyond the range of 1 to 50")
}else if(num == 20) {
  print("TRUE")
} else{
  print(num)
}
```
7. Change
At ISATU University’s traditional cafeteria, snacks can only be purchased with bills.
A long-standing rule at the concession stand is that snacks must be purchased with as few coins as possible.
There are three types of bills: 50 pesos, 100 pesos, 200 pesos, 500 pesos, 1000 pesos.

a. Write a function that prints the minimum number of bills that must be paid, given the price of the snack.
Input: Price of snack (a random number divisible by 50) Output: Minimum number of bills needed to purchase a snack.

```{r}
price <- as.numeric(readline(prompt = "Price of snack: "))
paste("Price of snack is", price)

bills <- c(50, 100, 200, 500, 1000)

min_bill <- (price %/% 50 + 1) * 50

count <- 0
for (bill in bills) {
  if (bill >= min_bill) {
    count <- bill
    break
  }
}
paste("Minimum bill needed to purchase a snack is", count)

```


  8. The following is each student’s math score for one semester. Based on this, answer the
following questions.
a. Create a dataframe from the above table. Write the R codes and its output.

```{r}
Name <- c("Annie", "Thea", "Steve", "Hanna")
Grade_1 <- c(85, 75, 65, 95)
Grade_2 <- c(65, 75, 55, 75)
Grade_3 <- c(85, 90, 80, 100)
Grade_4 <- c(100, 90, 85, 90)

data <- data.frame(Name, Grade_1, Grade_2, Grade_3, Grade_4)
data
```

b. Without using the rowMean function, output the average score of students whose average
math score over 90 points during the semester. write R code and its output.
Example Output: Annie’s average grade this semester is 88.75.

```{r}
  row_1 <- mean(as.numeric(data[1, 2:5]))
  paste(data[1, 1], "’s average grade this semester is", row_1)
  
  row_2 <- mean(as.numeric(data[2, 2:5]))
  paste(data[2, 1], "’s average grade this semester is", row_2)
  
  row_3 <- mean(as.numeric(data[3, 2:5]))
  paste(data[3, 1], "’s average grade this semester is", row_3)
  
  row_4 <- mean(as.numeric(data[4, 2:5]))
  paste(data[4, 1], "’s average grade this semester is", row_4)
  
```

c. Without using the mean function, output as follows for the tests in which the average score was less than 80 out of 4 tests.
Example output: The nth test was difficult.

```{r}
avg_grade_1 <- (85 + 75 + 65 + 95) / 4
avg_grade_2 <- (65 + 75 + 55 + 75) /4
avg_grade_3 <- (85 + 90 + 80 + 100) /4
avg_grade_4 <- (100 + 90 + 85 + 90) /4

avg_grade_1
avg_grade_2
avg_grade_3
avg_grade_4

if (avg_grade_1 < 80) {
  paste("The 1st test was difficult.")
}
 if (avg_grade_2 < 80) {
  paste("The 2nd test was difficult.")
}
 if (avg_grade_3 < 80) {
  paste("The 3rd test was difficult.")
}
 if (avg_grade_4 < 80) {
  paste("The 4th test was difficult.")
 } 
```

d. Without using the max function, output as follows for students whose highest score for a semester exceeds 90 points.
Example Output: Annie’s highest grade this semester is 95.

```{r}
annie <- c(85, 75, 65, 95)
highest_grade <- annie[1]
for (i in annie) {
  if (i > highest_grade) {
    highest_grade <- i
  }
} 
if(highest_grade > 90){
paste("Annie's highest grade this semester is", highest_grade)
}

thea <- c(65, 75, 55, 75)
highest_grade <- thea[1]
for(i in thea){
  if(i > highest_grade){
    highest_grade <- i
  }
}
if(highest_grade > 90){
paste("Thea's highest grade this semester is", highest_grade)
}

steve <- c(85, 90, 80, 100)
highest_grade <- steve[1]
for(i in steve){
  if(i > highest_grade){
    highest_grade <- i
  }
}
if(highest_grade > 90){
paste("Steve's highest grade this semester is", highest_grade)
}

hanna <- c(100, 90, 85, 90)
highest_grade <- hanna[1]
for(i in hanna){
  if(i > highest_grade){
    highest_grade <- i
  }
}
if(highest_grade > 90){
  paste("Hanna's highest grade this semester is", highest_grade)
}
```




