---
title: "RWorksheet_rocillo#4c"
author: "Cassandra Jielin Rocillo"
date: "2024-10-31"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

1. Use the dataset mpg
A data frame with 234 rows and 11 variables:
#’ \describe{
#’ \item{manufacturer}{manufacturer name}
#’ \item{model}{model name}
#’ \item{displ}{engine displacement, in litres}
#’ \item{year}{year of manufacture}
#’ \item{cyl}{number of cylinders}
#’ \item{trans}{type of transmission}
#’ \item{drv}{the type of drive train, where f = front-wheel drive, r = rear wheel drive, 4 = 4wd}
#’ \item{cty}{city miles per gallon}
#’ \item{hwy}{highway miles per gallon}
#’ \item{fl}{fuel type}
#’ \item{class}{"type" of car}
#’ }
"mpg"

• Download and open the mpg file. Upload it to your OWN environment
a. Show your solutions on how to import a csv file into the environment.

```{r}
library(readr)
imported_data <- read.csv("/cloud/project/worksheet/worksheet 4/worksheet4c/mpg.csv")
imported_data
```

b. Which variables from mpg dataset are categorical?
```{r}
cat_var <- names(imported_data)[sapply(imported_data, function(x) is.factor(x) || is.character(x))]

print("Categorical Variables:")
print(cat_var)
```

c. Which are continuous variables?

```{r}
cont_var <- names(imported_data)[sapply(imported_data, is.numeric)]
print("Continuous Variables:")
print(cont_var)
```

2. Which manufacturer has the most models in this data set? Which model has the most variations?
Show your answer.

```{r}
library(dplyr)
manufacturer <- imported_data %>% 
  group_by(manufacturer) %>%
  summarise(num_models = n_distinct(model)) %>%
  arrange(desc(num_models))

print("Manufacturer with the most models:")
manufacturer
```

a. Group the manufacturers and find the unique models. Show your codes and result.

```{r}
library(dplyr)

unique_models <- imported_data %>%
  group_by(manufacturer) %>%
  summarise(unique_models = list(unique(model)))

unique_models
```

b. Graph the result by using plot() and ggplot(). Write the codes and its result.

```{r}
library(ggplot2)
  
qplot(manufacturer, data = mpg,
geom = "bar",
fill = manufacturer)
```


2. Same dataset will be used. You are going to show the relationship of the model and the manufacturer.

```{r}

ggplot(mpg, aes(x = manufacturer, y = model, color = manufacturer)) +
  geom_point() +
  labs(title = "Manufacturer vs Model",
       x = "Manufacturer",
       y = "Model") 
```

a. What does ggplot(mpg, aes(model, manufacturer)) + geom_point() show?
  - The plot show a bar chart of the relationship of the model and the manufacturer.

b. For you, is it useful? If not, how could you modify the data to make it more informative?

  - The scatter plot is useful as it provides a basic overview of the model but there are overlapping points. Maybe we can use other code for the spacing to make it more wider so the text wont overlap to the other.

3. Plot the model and the year using ggplot(). Use only the top 20 observations. Write the codes and its results.

```{r}
 top_20 <- mpg[1:20, ]  

ggplot(top_20, aes(year,model)) +
  geom_point() +
  labs(title = "Top 20 Models by Year",
       x = "Year",
       y = "Model")
```

4. Using the pipe (%>%), group the model and get the number of cars per model. Show codes and its result

```{r}
model_counts <- mpg %>%
  group_by(model) %>%
  summarise(count = n()) 
model_counts
```

a. Plot using geom_bar() using the top 20 observations only.
The graphs shoudl have a title, labels and colors.
Show code and results.

```{r}
top_20_mpg <- mpg[1:20, ]

ggplot(top_20_mpg, aes(x = reorder(model, -year), fill = manufacturer)) +
  geom_bar() +
  labs(title = "Top 20 Cars by Model",
       x = "Model",
       y = "Count",
       fill = "Manufacturer")
```

b. Plot using the geom_bar() + coord_flip() just like what is shown below. Show codes and its result.

```{r}
ggplot(mpg, aes(x = model, fill = manufacturer)) +
  geom_bar(stat = "count") +  
  coord_flip() 
```


5. Plot the relationship between cyl - number of cylinders and displ - engine displacement using geom_point with aesthetic color = engine displacement. Title should be “Relationship between No. of Cylinders and Engine Displacement”.

```{r}
df_col <- mpg$drv
qplot(x = cyl,y = displ,
data = mpg,
geom = "point",
color = df_col,
main = "Relationship between No. of Cylinders(cyl) and Engine Displacement(displ)")

```

a. How would you describe its relationship? Show the codes and its result.
  - The output shows a scatter plot showing how the number of cylinders relates to engine displacement. 

6. Plot the relationship between displ (engine displacement) and hwy(highway miles per gallon).
Mapped it with a continuous variable you have identified in #1-c.
What is its result? Why it produced suchoutput?


```{r}
library(ggplot2)

ggplot(mpg, aes(x = displ, y = hwy, color = cty)) +
  geom_point() +
  labs(title = "Relationship between Engine Displacement and Highway MPG", 
       x = "Engine Displacement", 
       y = "Highway MPG")
```


6. Import the traffic.csv onto your R environment.

```{r}
traffic_data <- read.csv(file = "traffic.csv", sep = ",", header = T)
traffic_data
```

a. How many numbers of observation does it have? What are the variables of the traffic dataset the Show
your answer.

```{r}
observations <- nrow(traffic_data)
variables <- names(traffic_data)

paste("Number of Observations: ", observations)
paste("Variables of Traffic dataset: ", variables)
```

b. subset the traffic dataset into junctions. What is the R codes and its output?

```{r}

junctions_data <- subset(traffic_data, Junction == TRUE)  
junctions_data

```

c. Plot each junction in a using geom_line(). Show your solution and output.

```{r}

ggplot(junctions_data, aes(x = DateTime, y = Vehicles)) +
  geom_line() +
  labs(title = "Vehicle Count Over Time at Junction 1", x = "Date Time", y = "Number of Vehicles") +
  theme_minimal()

```

7. From alexa_file.xlsx, import it to your environment

```{r}
library(readxl)
alexa_data <- read_excel("alexa_file.xlsx")
alexa_data
```
a. How many observations does alexa_file has? What about the number of columns? Show your solution and answer.

```{r}
alexa_observations <- nrow(alexa_data)
alexa_column <- ncol(alexa_data)

paste("Number of Observations: ", alexa_observations)
paste("Number of columns: ", alexa_column)
```

b. group the variations and get the total of each variations. Use dplyr package. Show solution and answer.

```{r}
  variation_totals <- alexa_data %>%
  group_by(variation) %>%
  summarise(Count = n())

variation_totals
```


c. Plot the variations using the ggplot() function. What did you observe? Complete the details of the graph. Show solution and answer.

```{r}
qplot(variation, data = alexa_data,
geom = "bar",
fill = variation)
```


d. Plot a geom_line() with the date and the number of verified reviews. Complete the details of the graphs. Show your answer and solution.

```{r}
plt <- ggplot(alexa_data, mapping = aes(x = date, y = verified_reviews)) +
  geom_line(color = "blue") +
  theme_light() +
  xlab("Date") +
  ylab("Number of Verified Reviews") +
  ggtitle("Verified Reviews Over Time")

plt
```


e. Get the relationship of variations and ratings. Which variations got the most highest in rating? Plot a graph to show its relationship. Show your solution and answer.

```{r}
ggplot(alexa_data, aes(x = variation, y = rating)) +
  geom_point(color = "blue") +
  labs(title = "Variations vs Ratings",
       x = "Variation",
       y = "Ratings") 

```




