---
title: "RWorksheet_rocillo#2.Rmd"
author: "Cassandra Jielin Rocillo"
date: "2024-09-25"
output:
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

1. Createa vector using: operator
a. Sequence from -5 to 5. Write the R code and its output. Describe its output
```{r}
sequence <- (-5: 5)
print(sequence)
```
Describe the output.
  - The output shows number from-5 to 5

b. x <- 1:7. What will be the value of x?
```{r}
x <- 1:7
print(x)
```
2. create a vector using seq() function
a. seq(1, 3, by = 0.2)
```{r}
sequence <- seq(1, 3, by = 0.2)
print(sequence)
```
Describe the output.
  -Take output shows the number starting 1 to 3 in 0.2 sequence.

3. a factor has a census of its workers. There are 50 workers in total. the following list shows their ages.

```{r}
worker_age <- c(34, 28, 22, 36, 27, 18, 52, 39, 42, 29, 35, 31, 27, 22, 37, 34, 19, 20, 57, 49, 50, 37, 46, 25, 17, 37, 43, 53, 41, 51, 35, 24, 33, 41, 53, 40, 18, 44, 38, 41, 27, 39, 19, 30, 61, 54, 58, 26, 18)
```

a. Access the 3rd element, What is  the value?
```{r}
third <- worker_age[3]
print(third)
```
b. Access the 2nd and 4th element, what are the values?
```{r}
second <- worker_age[2]
fourth <- worker_age[4]
print(second)
print(fourth)
```

c. Access all but the 4th snd 12th element is not included.
```{r}
all <- worker_age[c(-4, -12)]
print(all)
```
4. Create a vector x <- c("first"=3, "second"=0, "third"=9).
Then named the vector, name(x).

```{r}
x <- c("first"=3, "second"=0, "third"=9)
print(x)
```
a. Print the result. Then access x[c("first", "third")]
```{r}
num <- x[c("first", "third")]
print(num)
```
Describe the output. 
   -The output shows how to name a vector aand how to access the elements using their names. 
   
5. Create a sequence x from -3:2.
a. Modify 2nd element and change it to 0:x[2] <- 0x

```{r}
x <- -3:2
x[2] <- 0
print(x)

```
Describe the output.
  -The output shows a number from _3 ro 2 but the second in the sequence results to zero as it was stated in the code the 2nd in the sequence of x will result to 0.
  
6. The following data shows the diesel fuel purchased by Mr. Cruz.

a. Create a data frame for month, price liter (php) and purchase-quality(liter). 
Write the R scripts and its output.

```{r}
month <- c("Jan", "Feb", "March", "Apr", "May", "June")
price_php <- c(52.50, 57.25, 60.00, 65.00, 74.25, 54.00)
liter <- c(25, 30, 40, 50, 10, 45)
purchased_data <- data.frame(Month = month, Price_php = price_php, Liter = liter)
print(purchased_data)

```

b. What is the average fuel expenditure of Mr. Cruz from Jan to June? 
Notes: Use weighted.mean(liter, purchase).
```{r}
average_exp <- weighted.mean(price_php, liter)
print(average_exp)
```
7. R has actually lots of built-in datasets. For example, the rivers data “gives the legths (in miles) of 141”major” rivers in North America, as compiled by the US Geological Survey”.
a. Type “rivers” in your R console. 
Create a vector data with 7 elements, containing the number of elements (length) in rivers, their sum (sum), mean (mean), median(median), variance(var), standard deviation(sd), minimum (min) and maximum (max).

```{r}
 data <- c(length(rivers), sum(rivers), mean(rivers), median(rivers), var(rivers),sd(rivers), min(rivers), max(rivers))
print(data)
```
b. What are the results? The result shows the:
  -Length of rivers - 141.0000
  -Sum of river lengths - 83, 357.0000
  -Mean of river lengths - 591.1844
  -Median of river lengths - 425.0000
  -var of river lengths - 243, 908.4086
  -sd of river lengths - 493.8708
  -min river length - 135.0000
  -max river length - 3710.0000

8. The table below gives the 25 most powerful celebrities and their annual pay as ranked by the editions of Forbes magazine and as listed on the Forbes.com website.
a. Create a vectors according to the above table.
```{r}
  celebrities <- c("Tom Cruise", "Rolling Stones", "Oprah Winfrey", "U2", "Tiger Woods", "Steven Spielberg", "Howard Stern", "50 Cent", "Cast of the Sopranos", "Dan Brown", " Bruce Springsteen", "Donald Trump", "Muhammad Ali", "Paul McCartney", "George Lucas", "Elton John", "David Letterman", "Phil Mickelson", "J.K Rowling", "Brad Pitt", "Peter Jackson", "Dr. Phil McGraw", "Jay Lenon", "Celine Dion", "Kobe Bryant")
pay <- c(67, 90, 225, 110, 90, 332, 302, 41, 52, 88, 55, 44, 55, 40, 233, 34, 40, 47, 75, 25, 39, 45, 32, 40, 31)

celebrity_data <- data.frame(Celebrities = celebrities, Pay = pay)
print(celebrity_data)
```
b. Modify the power ranking and pay of J.K. Rowling.
Change power ranking to 15 and pay to 90. 
Write the R scripts and its output.

```{r}
celebrity_data[celebrity_data$celebrities == "J.K. Rowling", "Pay"] <- 90
celebrity_data$Power_ranking <- 1:nrow(celebrity_data)
celebrity_data[celebrity_data$celebrities == "J.K. Rowling", "Power_ranking"] <- 15
print(celebrity_data)
```
c. Create an excel file from the table above and save it
as csv file(PowerRanking).  
Import the csv file into the RStudio. What is the R script?
```{r}
  write.csv(celebrity_data, "power_ranking.csv", row.names = FALSE)
import_data <- read.csv("power_ranking.csv")
print(import_data)
```
d. Access the rows 10 to 20 and save it as Ranks.RData.
Write the R script and its output.
```{r}
rank <- celebrity_data[10:20, ]
save(rank, file = "Ranks.RData")
print(rank)
```

9. Download the Hotels-Vienna https://tinyurl.com/Hotels-Vienna
a. Import the excel file into your RStudio.

what is the Rscript? 
```{r}
library(readxl)
hotel_data <- read_excel
print(head(hotel_data))
```
b. How many dimensions does the data set have?
What is the R script? What is its output?
```{r}
dimension <- dim(hotel_data)
print(dimension)
```
10. Create a list of ten (10) vegetables you ate during your lifetime. If none, just list down.
a. Write the R scripts and its output.
```{r}
vegetables <- c("cabbage", "spinach", "potato", "moringa", "green bean", "radish", "bottle gourd", "cucumber", "squash", "eggplant")
print(vegetables)
```
b. Add 2 additional vegetables after the last vegetables
in the list. 
```{r}
vegetables <- c(vegetables, "broccoli", "lettuce")
print(vegetables)
```
c. Add 4 additional vegetables after index 5. 
How many data points does your vegetable list have? 
```{r}
vegetables <- append(vegetables, c("onion", "garlic", "cauliflower", "carrots"), after = 5)
print(vegetables)

datapoints <- length(vegetables)
length(vegetables)
```
d. Remove the vegetables in index 5, 10, and 15. 
How many vegetables were left? Write the codes and its output.

```{r}
vegetables <- vegetables[-c(5, 10, 15)]
print(vegetables)

vegetables_left <- length(vegetables)
print(vegetables_left)

```

